---
title: "midterm"
author: "Sagar Jain"
date: "15 March 2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidytext)
library(wordcloud)
library(gutenbergr)
library(lubridate)
library(rvest)
library(stringr)
library(kableExtra)
library(magrittr)
library(tidyverse)
```

First, I download the dataset.

```{r prelim_1, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
tickets <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-12-03/tickets.csv")

head(tickets)
```

-----

#### Part a) 
##### _Question:_   


What is the most common violation and its average fine? What are the least common
violations and their respective fines?  

----

##### _Answer:_

----    


```{r part_1a, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
tail(names(sort(table(tickets$violation_desc))), 1)

Filtered_tickets_most <- tickets %>% filter(violation_desc == "METER EXPIRED CC") 

mean(Filtered_tickets_most$fine,na.rm = FALSE)



Filtered_tickets_least <- tickets %>% select(violation_desc,fine) %>% 
  filter(violation_desc %in% c("+4HR IN LOADING ZONE","INTERSECTION   CC","PARK-TROLLEY STOP 23","PK OVER TIME LIM 17","PRK MTR IMPROPER 11","STOP PROHIBITED  CC","STOPPED SAFE ZONE CC"))

kable(Filtered_tickets_least) %>% kable_styling()

```



----   

#### Part b) 
##### _Question:_

For the most common violation, create a plot in ggplot of the number of violations for
each month of 2017. What conclusions can you make?

----   

##### _Answer:_


----    

In the month of December there are less number of violations as compared to other months.
Number of violations keep on increasing every month from January to May, then there is a decrement in number from June to July and then it increased in August which is highest among all months.

```{r part_1b, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
n_distinct(month(Filtered_tickets_most$issue_datetime))


Filtered_tickets_most1 <- Filtered_tickets_most %>% mutate(month = month(issue_datetime))

Filtered_tickets_most1 <- Filtered_tickets_most1 %>% group_by(month) %>% summarise(count = n())
head(Filtered_tickets_most1)
ggplot(data = Filtered_tickets_most1)+
  geom_bar(mapping = aes(x=month,y=count),stat = "identity")+
  ggtitle("Monthly trend") + 
  xlab("Month") + 
  ylab("Count")
```


    
----   
#### Part c) 
##### _Question:_

The zip code for downtown Philadelphia is 19103. Are you more likely to get a ticket
here than elsewhere? To visualize this, create a graph of the number of violations for
each zip code 

----   



```{r part_1c, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
tickets_zipcode <- tickets %>% group_by(zip_code) %>% summarise(count = n()) %>% arrange(desc(count)) %>% na.omit()


ggplot(data = tickets_zipcode)+
  geom_bar(mapping = aes(x=zip_code,y=count),stat = "identity")+
  ggtitle("No. of violations distribution across different zipcode") + 
  xlab("Zipcode") + 
  ylab("No. of violations")

tickets_zipcode %>% select(zip_code,count) %>% filter(count ==max(count)) 

#since the number of violations is highest for zipcode 19103, therefore we are very likely to get violations in Philadelphia then anywhere else
```


----  



### Exercise 2   


-----

#### Part a) 
##### _Question:_   

Using the node .zn-body paragraph to extract the HTML content, create a data frame
which lists the number of cases and deaths for each state. In particular, create a data
frame with three columns, called state, cases, and deaths for all 51 states (including
the District of Columbia). For example, you will need to extract numeric values of 1 and
2, respectively, from the text 'including one death' and 'including two deaths'. Be sure
to make your code reproducible, as this is a live link that is being updated frequently.


----


```{r prelim_2a, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
covid19 <- "https://www.cnn.com/2020/03/03/health/us-coronavirus-cases-state-by-state/index.html" %>%
  read_html() %>% html_nodes(".zn-body__paragraph") %>% 
  html_text()


covid19.clean <- tibble(text = covid19[7:59])


covid19.clean <- covid19.clean %>% separate(text, c("State", "rest"), sep = ": ") %>%
  subset(!is.na(rest)) %>%
  mutate(State = str_trim(State))

covid19.onlystates <- covid19.clean[-c(46,40),]

covid19.onlystates.clean <- covid19.onlystates %>% separate(rest, c("Cases", "Deaths"), 
                                                            sep = "including ",extra = "merge") 

covid19.onlystates.clean <- covid19.onlystates.clean %>% separate(Cases, c("Cases", "unwanted"), sep = " ")

covid19.onlystates.clean <- covid19.onlystates.clean[,-3]

covid19.onlystates.clean <- covid19.onlystates.clean %>% separate(Deaths, c("Deaths", "unwanted"), sep = " ")

covid19.onlystates.clean <- covid19.onlystates.clean[,-4]

covid19.onlystates.clean$Deaths <- covid19.onlystates.clean$Deaths %>% 
                                  str_replace_all("one", "1") %>%
                                  str_replace_all("two", "2") %>%
                                  str_replace_all("three", "3")%>%
                                  str_replace_all("four", "4")%>%
                                  str_replace_all("five", "5")%>%
                                  str_replace_all("six", "6")%>%
                                  str_replace_all("seven", "7")%>%
                                  str_replace_all("eight", "8")%>%
                                  str_replace_all("nine", "9")%>%
                                  str_replace_all("ten", "10") 


covid19.onlystates.clean[is.na(covid19.onlystates.clean)] <- 0
head(covid19.onlystates.clean)

```


----    

#### Part b) 
##### _Question:_ 

Show the table in your RMarkdown file by using the function knitr::kable().



-----   

```{r prelim_2b, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
kable(covid19.onlystates.clean) %>%
  kable_styling()
```


----   

### Exercise 3    

----   


Taking advantage of the rvest package, turn the table at http://www.nature.com/articles/
ng.3097/tables/3 into an R data frame.

(a) Be sure to delete the rows "Genes with previous literature support (GRAIL)" and "New
genes without previous evidence" (and don't do it by using the row number).   




```{r prelim_3a, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
url <- 'http://www.nature.com/articles/ng.3097/tables/3'
table <- url %>% read_html() %>%  html_nodes("table") %>% html_table()
table <- table[[1]]
head(table)

#remove the excessive rows
table <- filter(table, Gene!="Genes with previous literature support (GRAIL)" & Gene != "New genes without previous evidence")
head(table)
```   

----   

#### part b

---
(b) Be sure to convert the p-value column to numbers.

```{r prelim_3b, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
#converting p-values to number by splitting the string apart, converting each value to number and then finally bringing it all together to put in the dataframe.
table_temp <- table %>% separate('Prioritization P value', c("val", "rest"), sep = "×")
table_temp <- table_temp %>% separate('rest', c("ten","pwr"), sep = "0")
table_temp$pwr <- parse_number(table_temp$pwr)
table_temp$pwr[is.na(table_temp$pwr)] <- 0
table_temp$val <- parse_number(table_temp$val)
table_temp <- table_temp %>% mutate('Prioritization P value'=val*(10^-pwr))
table <- table_temp[-c(4, 5, 6)]
head(table)
```

---

### 3(c)
The last column is shown in 3 rows in the journal, but most likely as one string in your
table. Use regular expressions to insert semicolons (i.e., ";") between each of the original
lines. For example,
"PI3K cascade (REACTOME, P = 6.2 x 10-13); Chronic myeloid leukemia (KEGG, P
= 1.6 x 10-12); Response to fibroblast growth factor stimulus (GO, P = 5.4 x 10-11)"

---
```{r prelim_3c, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
table$'Top-ranking reconstituted gene sets' <- table$'Top-ranking reconstituted gene sets' %>% str_replace_all( "[)]", ");") %>% substr(1,nchar(.)-1)
head(table)
```

   
---
###(d)
Show the table in your RMarkdown file by using the function knitr::kable().

---

```{r prelim_3d, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
kable(table, digits = 32) %>%  kable_styling()
```
