---
title: "Homework 5"
author: "Sagar Jain"
date: "March 2, 2020"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(magrittr)
library(dplyr)
library(broom)
```
----   

### Exercise 1    


----   


Downloaded the dataset.

```{r prelim_2, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
fed_rd <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-12/fed_r_d_spending.csv")
```   


#### Part a) 
##### _Question:_

Reformat this data with a separate variable called rd_budget_frac for the rd_budget as a fraction of total_outlays and two additional variables with upper and lower confidence bounds for rd_budget_frac from the linear model:
__rd_budget_frac ~ department + year__


----   

##### _Answer:_

At First, I created the variable __rd_budget_frac__.   
Then, I did the linear model, and after that, I calculate the confidence intervals.   
Finally, I left join them with the original dataset.   

----    



```{r part_2a, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
fed_rd.new <- fed_rd %>%
  mutate(rd_budget_frac = rd_budget/total_outlays)
fit <- lm(rd_budget_frac ~ department + year, fed_rd.new)
conf.int <- augment(fit) %>%
  mutate(fit = .fitted) %>%
  mutate(upper = .fitted + 1.96 * .se.fit) %>%
  mutate(lower = .fitted + 1.96 * .se.fit) %>%
  select(rd_budget_frac, department, year, fit, upper, lower)
fed_rd.new <- fed_rd.new %>%
  left_join(.,conf.int)
  
head(fed_rd.new)
```


---   

#### Part b) 
##### _Question:_
Create four plots showing rd_budget_frac (along with the upper and lower confidence bounds from (a)) as a function of year for NASA, NSF, DHS, and DOD. (Be sure your figure looks polished.) Comment on any patterns you find.

----   

##### _Answer:_
I created a function to plot all 4 plots, since it would be repeating lines of codes. Then, I just called the function foreach department.      
    
----    



```{r part_2b, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
ploting <- function(department){
  
  coefs.plot=fed_rd.new[which(fed_rd.new$department==department),]
  
  ggplot(coefs.plot, aes(year, rd_budget_frac))+
    geom_point()+
    geom_line(aes(y=lower), color = "red", linetype = "dashed")+
    geom_line(aes(y=upper), color = "red", linetype = "dashed")+
    geom_smooth()
}
ploting("NASA")
ploting("NSF")
ploting("DHS")
ploting("DOD")
```

**_Comments:_**
Except for _DHS_, the rest have a _descending relation_ through the years, _DHS_ is a _flat horizontal line until 2001_, approximately.   



----

### Exercise 3    

-----

Table 16 of the file UN MigrantStockByOriginAndDestination 2015.xlsx (in the Week 5 Canvas folder) shows migration from one country to another in 2015. By eliminating the rows and columns that don't correspond to countries and then converting to a tidy dataset using gather() (i.e., a data frame with three columns: one for country of origin, one for country of destination, and the third indicating the number of people who immigrated from one country to the other), find

----   


First, I upload the file and clean the dataset.      
    
    
----    





```{r prelim_3, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
getwd()

untidy_migrant <- readxl::read_excel("UN_MigrantStockByOriginAndDestination_2015.xlsx", sheet = "Table 16", skip = 15)
head(untidy_migrant)

untidy_migrant <- untidy_migrant %>%
  rename(to_country = ...2)

untidy_migrant <- untidy_migrant %>%
  rename(country_code = ...4)

untidy_migrant <- filter(untidy_migrant,country_code < 900)
head(untidy_migrant)
```   
    
    
    
---   

#### Part a) 
##### _Question:_   

The top five countries from which people migrate to Canada.    


----   

##### _Answer:_    


After cleaning the dataset, is easy to use _gather()_ to get the answer to this question.
 
 
 
 

```{r part_3a, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
migrate_to_canada <- untidy_migrant[which(untidy_migrant$to_country == "Canada"),]
migrate_to_canada <- migrate_to_canada %>% gather(`Afghanistan`:`Zimbabwe`, key = "from_country", value = "count")

migrate_to_canada_cleaned <- select(migrate_to_canada,c(to_country,
                           from_country,count))

migrate_to_canada_cleaned %>%
  arrange(desc(count)) %>%
  head(5)

```    



#### Part b) 
##### _Question:_   

The top five countries to which people migrate from Canada.   


----   

##### _Answer:_    


Using _gather()_ again, I counted the countries from Canada.   
   
    
----    


```{r part_3b, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
migrate_from_canada <- untidy_migrant %>%filter(country_code < 900)
head(arrange(migrate_from_canada, desc(Canada)), n = 5)$to_country
```   



#### Part c) 
##### _Question:_   

The top 10 migration pairs of countries.   


----   

##### _Answer:_    


First, I keep only the countries and exclude regions.   
Then, I just looked for the top 10 pairs.       
    
    
----    

  

```{r part_3c, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
migration_pairs <- untidy_migrant %>%
  gather(`Afghanistan`:`Zimbabwe`, key = "from_country", value = "count") %>% arrange(desc(count))
#Excluding the unnecessary columns
drop_columns <- c("Total","Other North", "Other South", "Country_code", "..1", "..3", "..5")
pairs_top_10 <- migration_pairs[ , !(names(migration_pairs) %in% drop_columns)]
pairs_top_10 %>%
  select(to_country, from_country) %>%
  head(10)
```   